---
title: "PA1_template.RMD"
author: "Daniel Castineyra"
date: "26 de noviembre de 2017"
output: html_document
---
1. Loading the data and taking out the NA values
```{r, echo=TRUE}
setwd("~/Data Science/Rep Research Assig1")
datos<-read.table("activity.csv",sep = ",",header = TRUE)
activity<-subset(datos,!is.na(datos$steps))
library(dplyr)
```

2. Histogram of the total number of steps taken each day

```{r, echo=TRUE}
## Creating a data frame with the total steps by day
day<-unique(activity$date)
day<-as.Date(day)
total.steps<-as.vector(tapply(activity$steps,activity$date,sum))
total.steps<-subset(total.steps,!is.na(total.steps))
## Plotting the histogram
plot(day,total.steps,type="h",xlab="Day",ylab="Total Steps",main="Total Steps by Day")
```

3. Mean and median number of steps taken each day

```{r, echo=TRUE}
## Average and Median by day
Average.Steps = as.vector(tapply(activity$steps,activity$date,mean))
Average.Steps<-subset(Average.Steps,!is.na(Average.Steps))
Median.Steps = as.vector(tapply(activity$steps,activity$date,median))
Median.Steps<-subset(Median.Steps,!is.na(Median.Steps))
mm_by_day<-data.frame(Day=day,Average.Steps = Average.Steps,Median.Steps = Median.Steps)
```

This table shows the Average and Median of Steps per Day

```{r,echo=TRUE}
mm_by_day
```

4. Time series plot of the average numbers of steps taken
```{r, echo=TRUE}
## Time Series of Average Steps
Average.perinterval<-as.vector(tapply(activity$steps,activity$interval,mean))
interval<-unique(activity$interval)
plot(interval,Average.perinterval,type="l",xlab = "Interval",ylab = "Average Steps",main = "Average Steps by Interval")
```

5. The 5-minute interval that, on average, containis the maximum number of steps
``` {r, echo=TRUE}
## Interval with the maximum average steps
max.interval<-interval[which.max(Average.perinterval)]
```

Maximum is `r max.interval`

6. Code to describe and show a strategy for imputing missing data
To fill the NA values we use the average of each interval
```{r,echo=TRUE}
## Number of missing values
mv<-subset(datos,is.na(datos$steps)==TRUE)
nmv<-length(mv$steps)
## Filling missing values with average of the interval
Average.perinterval2<-c(Average.perinterval,Average.perinterval,Average.perinterval,Average.perinterval,Average.perinterval,Average.perinterval,Average.perinterval,Average.perinterval)
mv<-mutate(mv,steps=Average.perinterval2)
##new data
activity2<-merge(activity,mv,all=TRUE)
day2<-unique(activity2$date)
day2<-as.Date(day2)
total.steps2<-as.vector(tapply(activity2$steps,activity2$date,sum))
```

7. Histogram of the total number of steps taken each day after missing values are imputed
```{r,echo=TRUE}
## Plotting the histogram
plot(day2,total.steps2,type="h",xlab="Day",ylab="Total Steps",main="Total Steps by Day")
## Average and Median by day
Average.Steps2 = as.vector(tapply(activity2$steps,activity2$date,mean))
Median.Steps2 = as.vector(tapply(activity2$steps,activity2$date,median))
mm_by_day2<-data.frame(Day=day2,Average.Steps = Average.Steps2,Median.Steps = Median.Steps2)
```

8. Panel plot comparting the average number of steps taken per 5-minute interval across weekdays and weekends
```{r,echo=TRUE}
activity3<-mutate(activity2,day=weekdays(as.Date(activity2$date)))
weekday<-subset(activity3,activity3$day=="lunes" | activity3$day=="martes" | activity3$day == "miércoles" | activity3$day=="jueves")
weekend<-subset(activity3,activity3$day== "viernes" | activity3$day=="sábado" | activity3$day=="domingo")
Average.weekday<-as.vector(tapply(weekday$steps,weekday$interval,mean))
Average.weekend<-as.vector(tapply(weekend$steps,weekend$interval,mean))
interval1<-unique(activity3$interval)
par(mfrow=c(2,1))
plot(interval1,Average.weekday,type="l",xlab = "Interval",ylab = "Average Weekday",main = "Average Steps by Interval")
plot(interval1,Average.weekend,type="l",xlab = "Interval",ylab = "Average Weekend",main = "Average Steps by Interval")
````
